\documentclass[10pt]{article} 
\usepackage[T1]{fontenc} 
\usepackage[utf8]{inputenc}
\usepackage{geometry} 
\geometry{verbose,marginparwidth=0.5in,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in} 
\usepackage{lmodern}

\usepackage{booktabs}

\usepackage{enumitem}
% \setlist{nosep}

% \usepackage{amsfonts}
% \usepackage{amsmath}
\usepackage{comment}

\usepackage{mathtools}
\usepackage{bbm}
\newcommand{\one}{\mathbbm{1}}



\begin{document}


\begin{center}
\textbf{Econ 512}

\emph{Fall 2018}\\[1em]

Homework 1 -- Introduction/Review of Matlab Basics\\
Due 9/5/2018\\[3em]
\end{center}

\noindent
Write and execute a matlab program (m-file) that will make and print the following calculations. Turn in your program and file containing theresults. Read about the ``diary'' command for producing your results file. Please carefully comment your files so I can easily read your answers.

\bigskip 

\noindent
\textbf{1.} Given the following vector of values $X=[1,1.5,3,4,5,7,9,10]$
construct the values of the functions  

$$Y1=-2+.5X$$ 

and 

$$Y2=-2+.5X^{2}$$ 

and plot $Y1$and $Y2$ against $X$ in a single graph. 

\bigskip 

\noindent
\textbf{2.} Create a 200x1 vector $X$ containing evenly-spaced numbers between [-10, 20] and calculate the sum of the elements of the vector. 

\bigskip 

\noindent
\textbf{3.} Given a matrix $A$=$%
\begin{bmatrix}
2 & 4 & 6 \\ 
1 & 7 & 5 \\ 
3 & 12 & 4%
\end{bmatrix}%
$ and vector $b$=$%
\begin{bmatrix}
-2 \\ 
3 \\ 
10%
\end{bmatrix}%
.$  Calculate $C=A^{\prime }b$, \ $D=(A^{\prime }A)^{-1}b,$  $%
E=\sum_{i}\sum_{j}[a_{ij}b_{i}],$ and $F=$ matrix $A$ with the 2nd row and
3rd column deleted. \ Solve the system of linear equations $Ax=b$ for the
vector $x.$

\noindent
\textbf{4.} Using the matrix $A$ in problem 3, create a 15x15 block diagonal
matrix $B$ =  $%
\begin{bmatrix}
A & 0 & 0 & 0 & 0 \\ 
0 & A & 0 & 0 & 0 \\ 
0 & 0 & A & 0 & 0 \\ 
0 & 0 & 0 & A & 0 \\ 
0 & 0 & 0 & 0 & A%
\end{bmatrix}%
$\ where 0 is a 3x3 matrix of zeros.

\bigskip 

\noindent
\textbf{5.} Create a 5x3 matrix of random draws from a normal distribution with
mean= 10 and standard deviation=5. \ \ Call it $A$. \ Convert $A$ to a
matrix of zeros and ones where the element of the new matrix is 0 if $%
a_{ij}<10$ and 1 if $a_{ij}\geq 10.$ 

\bigskip 

\noindent
\textbf{6.}  In the github repository {\tt Homework1} for the class, you will find the file {\tt datahw1.csv} that is a
comma-delimited spreadsheet file that you will use to answer this question.
\ The dataset constrains 4392 firm/year observations for Taiwanese
manufacturing firms. \ The columns of the data matrix are (in order): \ (1)
firm id, (2) year, (3) dummy variable equal to one if the firm is an
exporter (Export), (4) dummy variable equal to one if the firm conducts R\&D
(RD), (5) productivity index (prod), (6) capital stock (cap). \ \ Construct
the OLS estimator for the
regression equation: \ 

$$prod_{it}=\beta _{0}+\beta _{1}Export_{it}+\beta _{2}RD_{it}+\beta
_{3}cap_{it} + \varepsilon _{it}$$
Report the point estimates and standard errors.


\end{document}
